### 2013-07-30 Tex/chapters/methods.tex
Ergänzungen zur autokorrelationszeit und dynamischem exponent

        \(\frac{\chi(t)}{\chi(0)}\), which should decay exponentially

@@

        \(\tau\) on the system size \(L\) is then given by \(\tau \propto L^z\).

@@

        section \ref{sec:results}.

>>>>

        \(\frac{\chi(t)}{\chi(0)}\) with
        \begin{equation}
            \chi(t) = \int \de t' [m(t') -\avg{m}][m(t'+t)-\avg{m}]
        \end{equation}
        which should decay exponentially

@@

        \(\tau\) on the system size \(L\) is at \(T_{c}\) given by \(\tau \propto L^z\).
        More general the power law \(\tau \propto \xi^z\) holds true, where
        \(\xi\) is the \emph{correlation length}, which diverges at
        \(T_{c}\) and is then limited by the size of the simulated lattice.

@@

        section \ref{ssec:results:autocorr}. Also according to \cite{NewmanBarkema1999}
        \(z\) is independent of the lattice structure, which ensures that
        the simulation will benefit from the Wolff cluster algorithm at
        any \(\sigma\).

### 2013-07-30 Tex/chapters/results.tex
Ergänzungen zur autokorrelationszeit und dynamischem exponent

\subsection{Short Analysis of the Autocorrelation Time  $\tau$}
    % Sollte das hier rein?
    Before the results are presented, a short analysis of the
    autocorrelation time \(\tau\) is given, to illustrate the benefits
    of the used Wolff cluster algorithm.\\
    In \ref{fig:autocorr}\subref{fig:autocorr:temperatures}
    the autocorrelation time for the magnetization per site \(m\) at \(\sigma=0\) is plotted.
    \begin{figure}[htbp]
        \centering
        \subfigure[in Dependence on the Temperature $T$][]
        {
            \label{fig:autocorr:temperatures}
            \includegraphics[width=0.45\textwidth]{plots/autocorrT}
        }
        \subfigure[in Dependence on the System Size $L$][]
        {
            \label{fig:autocorr:powerLaw}
            \includegraphics[width=0.45\textwidth]{plots/autocorrL}
        }
        \caption[The Autocorrelation Time $\tau$]
        {
            Dependece of the autocorrelation time $\tau$ on
            \subref{fig:autocorr:temperatures} the temperature $T$ for
                $\sigma=0$ and
            \subref{fig:autocorr:powerLaw} the system size $L$ (errorbars
                are the standard error determined by bootstrapping)
        }
        \label{fig:autocorr}
    \end{figure}
    The plateau at low temperatures is easy to understand considering the
    effects of the Wolff algorithm. At low temperatures it flips nearly every
    spin in every step, thus the correlation drops to zero after one sweep.
    Alternatively the parallel tempering algorithm swaps spin configurations
    with possibly different signs, thus having the same effect.
    Also note that the maximum of \(\tau\) is not at \(T_c\) but at a higher
    temperature. The cause is probably the effectiveness of the Wolff cluster
    algorithm at \(T_c\).
    Obviously the autocorrelation time \(\tau\) increases with the system
    size. In fact it obeys a power law, which is the expected behaivior of
    a dynamical critical exponent \(z\) as mentioned in section \ref{sssec:eqtime}.
    In fig. \ref{fig:autocorr}\subref{fig:autocorr:powerLaw} \(\tau\) is plotted
    over \(L\) and the dynamical exponent \(z'\) is determined as \(z'=0.64(2)\).
    This exponent is much higher than the one of Wolff cluster
    algorithm at criticality, which is \(z=0.25\)\cite{NewmanBarkema1999}.
    So it represents the dynamical exponent for parallel tempering with
    Metropolis above criticality and thus is not exactly a critical exponent.
    Nevertheless it is interesting, that it obeys a power law just the same.
    Anyhow it is a proof for the effectiveness of the Wolff cluster algorithm
    at criticality.

>>>>

\subsection{Short Analysis of the Autocorrelation Time  $\tau$}
\label{ssec:results:autocorr}
    % Sollte das hier rein?
    Before the results are presented, a short analysis of the
    autocorrelation time \(\tau\) is given, to illustrate the benefits
    of the used Wolff cluster algorithm.\\
    In \ref{fig:autocorr}\subref{fig:autocorr:temperatures}
    the autocorrelation time for the magnetization per site \(m\) at \(\sigma=0\) is plotted.
    \begin{figure}[htbp]
        \centering
        \subfigure[in Dependence on the Temperature $T$][]
        {
            \label{fig:autocorr:temperatures}
            \includegraphics[width=0.45\textwidth]{plots/autocorrT}
        }
        \subfigure[in Dependence on the System Size $L$][]
        {
            \label{fig:autocorr:powerLaw}
            \includegraphics[width=0.45\textwidth]{plots/autocorrL}
        }
        \caption[The Autocorrelation Time $\tau$]
        {
            Dependece of the autocorrelation time $\tau$ on
            \subref{fig:autocorr:temperatures} the temperature $T$ for
                $\sigma=0$ and
            \subref{fig:autocorr:powerLaw} the system size $L$ (errorbars
                are the standard error determined by bootstrapping)
        }
        \label{fig:autocorr}
    \end{figure}
    The plateau at low temperatures is easy to understand considering the
    effects of the Wolff algorithm. At low temperatures it flips nearly every
    spin in every step, thus the correlation drops to zero after one sweep.
    Alternatively the parallel tempering algorithm swaps spin configurations
    with possibly different signs, thus having the same effect.
    Also note that the maximum of \(\tau\) is not at \(T_c\) but at a higher
    temperature. The cause is probably the effectiveness of the Wolff cluster
    algorithm at \(T_c\), hence \(\tau \approx 1\) at \(T_{c}\).
    Obviously the autocorrelation time \(\tau\) increases with the system
    size. In fact it obeys a power law \(\tau \propto \L^z\), which is
    the expected behavior of a dynamical exponent \(z\) as mentioned in
    section \ref{sssec:eqtime}.
    In fig. \ref{fig:autocorr}\subref{fig:autocorr:powerLaw} \(\tau\) is plotted
    over \(L\). Note that the exponent \(z'=0.64(2)\) determined by this plot is
    not comparable to the known critical exponent \(z=0.25\) \cite{NewmanBarkema1999}
    for the Wolff cluster algorithm. On the one hand the number of sweeps
    in which \(\tau\) is measured in this thesis are more than a standard
    Metropolis sweep to which \(z\) normally corresponds, thus \(z'\) would
    need to be normalized by some factor \(>1\), because each sweep
    alongside the \(N\) Metropolis flips also a cluster of \(\ge 1\) spin is
    flipped. On the other hand the \(\tau\) used for the fit are not the
    \(\tau\) at \(T_{c}\) but the maximum \(\tau\) of all \(T\).
    Nevertheless it is interesting, that it obeys a power law
    \(\tau \propto L^{z'}\) just the same. This suggests, that \(\xi \gtrsim L\)
    is satisfied at the temperature of maximum \(\tau\) and that \(z'\) is a
    valid upper bound of \(z\).
    This proofs the effectiveness of the Wolff cluster algorithm at
    criticality.\\
    This is of course only a rough analysis of existing data which was
    generated for another purpose. For a more in depth inspection, one
    would perform the simulation at the critical point, which will be
    determined in the next sections.
